---
title: "John.RMD"
author: "John Ivanov"
date: "2022-07-25"
output: html_document
---
```{r, , echo=FALSE}
alz <- read.csv("alzheimer_data.csv")
attach(alz)
library(car)
library(ggplot2)
```
## Scatter plot matrix of question data
```{r, echo=FALSE}
 
ran <- sample(1:2700,2700)
new_data <- alz[ran,]

pairs(~diagnosis +shopping+bills+taxes+stove+mealprep+travel, data = new_data, main="Simple Scatterplot Matrix")
```

## Barplots relating diagnosis with certain question variables
```{r, echo=FALSE}

shop_graph = ggplot(data = alz, aes(x = as.factor(shopping), fill = as.factor(diagnosis))) + geom_bar()
shop_graph

bills_graph = ggplot(data = alz, aes(x = as.factor(bills), fill = as.factor(diagnosis))) + geom_bar()
bills_graph

taxes_graph = ggplot(data = alz, aes(x = as.factor(taxes), fill = as.factor(diagnosis))) + geom_bar()
taxes_graph

stove_graph = ggplot(data = alz, aes(x = as.factor(stove), fill = as.factor(diagnosis))) + geom_bar()
stove_graph

mealprep_graph = ggplot(data = alz, aes(x = as.factor(mealprep), fill = as.factor(diagnosis))) + geom_bar()
mealprep_graph

travel_graph =  ggplot(data = alz, aes(x = as.factor(travel), fill = as.factor(diagnosis))) + geom_bar()
travel_graph

chess_graph =  ggplot(data = alz, aes(x = as.factor(games), fill = as.factor(diagnosis))) + geom_bar()
chess_graph
```

## Chi-Squared Test for each question variable
```{r, echo=FALSE}
shop_table = table(diagnosis,shopping)
shop_chisquaretest = chisq.test(shop_table)
shop_chisquaretest
shop_p_val = shop_chisquaretest$p.value
```
Because the p-value, `r shop_p_val`, is less than 0.05, we can conclude that there is a relationship between shopping and diagnosis.


```{r, echo=FALSE}
bills_table = table(diagnosis,bills)
bills_chisquaretest = chisq.test(bills_table)
bills_chisquaretest
bills_p_val = bills_chisquaretest$p.value
```
Because the p-value, `r bills_p_val`, is less than 0.05, we can conclude that there is a relationship between bills and diagnosis.


```{r, echo=FALSE}
taxes_table = table(diagnosis,taxes)
taxes_chisquaretest = chisq.test(taxes_table)
taxes_chisquaretest
taxes_p_val = taxes_chisquaretest$p.value
```
Because the p-value, `r taxes_p_val`, is less than 0.05, we can conclude that there is a relationship between taxes and diagnosis.


```{r, echo=FALSE}
stove_table = table(diagnosis,stove)
stove_chisquaretest = chisq.test(stove_table)
stove_chisquaretest
stove_p_val = stove_chisquaretest$p.value
```
Because the p-value, `r stove_p_val`, is less than 0.05, we can conclude that there is a relationship between stove and diagnosis.


```{r, echo=FALSE}
meal_table = table(diagnosis,mealprep)
meal_chisquaretest = chisq.test(meal_table)
meal_chisquaretest
meal_p_val = meal_chisquaretest$p.value
```
Because the p-value, `r meal_p_val`, is less than 0.05, we can conclude that there is a relationship between meal and diagnosis.


```{r, echo=FALSE}
travel_table = table(diagnosis,travel)
travel_chisquaretest = chisq.test(travel_table)
travel_chisquaretest
travel_p_val = travel_chisquaretest$p.value
```
Because the p-value, `r travel_p_val`, is less than 0.05, we can conclude that there is a relationship between travel and diagnosis.


```{r, echo=FALSE}
chess_table = table(diagnosis,games)
chess_chisquaretest = chisq.test(chess_table)
chess_chisquaretest
chess_p_val = chess_chisquaretest$p.value
```
Because the p-value, `r chess_p_val`, is less than 0.05, we can conclude that there is a relationship between chess and diagnosis.


