---
title: "John.RMD"
author: "John Ivanov"
date: "2022-07-25"
output: html_document
---
```{r, , echo=FALSE}
alz <- read.csv("alzheimer_data.csv")

attach(alz)
library(car)
library(ggplot2)
```
## Scatter plot matrix of question data
```{r, echo=FALSE}
 
ran <- sample(1:2700,2700)
new_data <- alz[ran,]

pairs(~diagnosis +shopping+bills+taxes+stove+mealprep+travel, data = new_data, main="Simple Scatterplot Matrix")
```

## Barplots relating diagnosis with certain question variables
```{r, echo=FALSE}


shop_graph = ggplot(data = alz, aes(x = as.factor(shopping), fill = as.factor(diagnosis))) + geom_bar()
shop_graph

bills_graph = ggplot(data = alz, aes(x = as.factor(bills), fill = as.factor(diagnosis))) + geom_bar()
bills_graph

taxes_graph = ggplot(data = alz, aes(x = as.factor(taxes), fill = as.factor(diagnosis))) + geom_bar()
taxes_graph

stove_graph = ggplot(data = alz, aes(x = as.factor(stove), fill = as.factor(diagnosis))) + geom_bar()
stove_graph

mealprep_graph = ggplot(data = alz, aes(x = as.factor(mealprep), fill = as.factor(diagnosis))) + geom_bar()
mealprep_graph

travel_graph =  ggplot(data = alz, aes(x = as.factor(travel), fill = as.factor(diagnosis))) + geom_bar()
travel_graph

alz <- alz |>
  filter(games != 8)
chess_graph =  ggplot(data = alz, aes(x = as.factor(games), fill = as.factor(diagnosis))) + labs(fill = "Diagnosis", title = "Amount of Difficulty Subject Faced Performing a Game or Hobby based on Alzheimer Diagnosis", x = "Difficulty Level", y = "Number of Patients")+geom_bar()+ theme(plot.title = element_text(size = 10))
chess_graph



trailA_graph =  ggplot(data = alz, aes(x = as.factor(diagnosis), y=traila)) +labs(title = "Amount of Time Taken to Complete a Trail Making Task Based on Alzheimer Diagnosis", x = "Diagnosis", y = "Time Taken (seconds)")+ geom_boxplot(fill = "Purple", color = "black")+ theme(plot.title = element_text(size = 13))
trailA_graph


trailB_graph =  ggplot(data = alz, aes(x = as.factor(diagnosis), y=trailb)) + geom_boxplot()
trailB_graph

hall_graph =  ggplot(data = alz, aes(x = as.factor(hallsev), fill = as.factor(diagnosis))) + geom_bar()
hall_graph

anx_graph =  ggplot(data = alz, aes(x = as.factor(anxsev), fill = as.factor(diagnosis))) + geom_bar()
anx_graph

depress_graph =  ggplot(data = alz, aes(x = as.factor(depdsev), fill = as.factor(diagnosis))) + geom_bar()
depress_graph

gend_graph =  ggplot(data = alz, aes(x = as.factor(female), fill = as.factor(diagnosis))) + geom_bar()
gend_graph

weight_graph =  ggplot(data = alz, aes(x = as.factor(diagnosis), y=weight)) + geom_boxplot()
weight_graph

hrate_graph =  ggplot(data = alz, aes(x = as.factor(diagnosis), y=hrate)) + geom_boxplot()
hrate_graph

blood_graph =  ggplot(data = alz, aes(x = as.factor(diagnosis), y=bpsys)) + geom_boxplot()
blood_graph

motor_graph =  ggplot(data = alz, aes(x = as.factor(motsev), fill = as.factor(diagnosis))) + geom_bar()
motor_graph
```

## Chi-Squared Test for each question variable
```{r, echo=FALSE}
shop_table = table(diagnosis,shopping)
shop_chisquaretest = chisq.test(shop_table)
shop_chisquaretest
shop_p_val = shop_chisquaretest$p.value
```
Because the p-value, `r shop_p_val`, is less than 0.05, we can conclude that there is a relationship between shopping and diagnosis.


```{r, echo=FALSE}
bills_table = table(diagnosis,bills)
bills_chisquaretest = chisq.test(bills_table)
bills_chisquaretest
bills_p_val = bills_chisquaretest$p.value
```
Because the p-value, `r bills_p_val`, is less than 0.05, we can conclude that there is a relationship between bills and diagnosis.


```{r, echo=FALSE}
taxes_table = table(diagnosis,taxes)
taxes_chisquaretest = chisq.test(taxes_table)
taxes_chisquaretest
taxes_p_val = taxes_chisquaretest$p.value
```
Because the p-value, `r taxes_p_val`, is less than 0.05, we can conclude that there is a relationship between taxes and diagnosis.


```{r, echo=FALSE}
stove_table = table(diagnosis,stove)
stove_chisquaretest = chisq.test(stove_table)
stove_chisquaretest
stove_p_val = stove_chisquaretest$p.value
```
Because the p-value, `r stove_p_val`, is less than 0.05, we can conclude that there is a relationship between stove and diagnosis.


```{r, echo=FALSE}
meal_table = table(diagnosis,mealprep)
meal_chisquaretest = chisq.test(meal_table)
meal_chisquaretest
meal_p_val = meal_chisquaretest$p.value
```
Because the p-value, `r meal_p_val`, is less than 0.05, we can conclude that there is a relationship between meal and diagnosis.


```{r, echo=FALSE}
travel_table = table(diagnosis,travel)
travel_chisquaretest = chisq.test(travel_table)
travel_chisquaretest
travel_p_val = travel_chisquaretest$p.value
```
Because the p-value, `r travel_p_val`, is less than 0.05, we can conclude that there is a relationship between travel and diagnosis.


```{r, echo=FALSE}
chess_table = table(diagnosis,games)
chess_chisquaretest = chisq.test(chess_table)
chess_chisquaretest
chess_p_val = chess_chisquaretest$p.value
```
Because the p-value, `r chess_p_val`, is less than 0.05, we can conclude that there is a relationship between chess and diagnosis.

```{r, echo=FALSE}
tA_tTest = t.test(traila~bin_diag)
tA_tTest
tA_p_val = tA_tTest$p.value
```
Because the p-value, `r tA_p_val`, is less than 0.05, we can conclude that there is a relationship between trialA and diagnosis.

```{r, echo=FALSE}
tB_tTest = t.test(trailb~bin_diag)
tB_tTest
tB_p_val = tB_tTest$p.value
```
Because the p-value, `r tB_p_val`, is less than 0.05, we can conclude that there is a relationship between trailB and diagnosis.

```{r, echo=FALSE}
hall_table = table(diagnosis,hallsev)
hall_chisquaretest = chisq.test(hall_table)
hall_chisquaretest
hall_p_val = hall_chisquaretest$p.value
```
Because the p-value, `r hall_p_val`, is less than 0.05, we can conclude that there is a relationship between hallucinations and diagnosis.

```{r, echo=FALSE}
anx_table = table(diagnosis,anxsev)
anx_chisquaretest = chisq.test(anx_table)
anx_chisquaretest
anx_p_val = anx_chisquaretest$p.value
```
Because the p-value, `r anx_p_val`, is less than 0.05, we can conclude that there is a relationship between anxiety and diagnosis.

```{r, echo=FALSE}
depress_table = table(diagnosis,depdsev)
depress_chisquaretest = chisq.test(depress_table)
depress_chisquaretest
depress_p_val = depress_chisquaretest$p.value
```
Because the p-value, `r depress_p_val`, is less than 0.05, we can conclude that there is a relationship between depression and diagnosis.

```{r, echo=FALSE}
gend_table = table(diagnosis,female)
gend_chisquaretest = chisq.test(gend_table)
gend_chisquaretest
gend_p_val = gend_chisquaretest$p.value
```
Because the p-value, `r gend_p_val`, is less than 0.05, we can conclude that there is a relationship between gender and diagnosis.

```{r, echo=FALSE}
weight_tTest = t.test(weight~bin_diag)
weight_tTest
weight_p_val = weight_tTest$p.value
```
Because the p-value, `r weight_p_val`, is less than 0.05, we can conclude that there is a relationship between weight and diagnosis.

```{r, echo=FALSE}

hrate_tTest = t.test(hrate~bin_diag)
hrate_tTest
hrate_p_val = hrate_tTest$p.value
```
Because the p-value, `r hrate_p_val`, is less than 0.426, we cannot conclude that there is a relationship between hearth rate and diagnosis.

```{r, echo=FALSE}

blood_tTest = t.test(bpsys~bin_diag)
blood_tTest
blood_p_val = blood_tTest$p.value
```
Because the p-value, `r blood_p_val`, is less than 0.05, we can conclude that there is a relationship between blood pressure and diagnosis.

```{r, echo=FALSE}
motor_table = table(diagnosis,motsev)
motor_chisquaretest = chisq.test(motor_table)
motor_chisquaretest
motor_p_val = motor_chisquaretest$p.value
```
Because the p-value, `r motor_p_val`, is less than 0.05, we can conclude that there is a relationship between motor disturbance and diagnosis.